# COURSERA - Reproducible Research - Peer Assessment 1

```{r}
library(lattice)
```

## Loading and preprocessing the data
Load activity.csv  
Convert dates to R Date class
```{r}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date,"%Y-%m-%d")
```

## What is the mean total number of steps taken per day?
Computation of the total number of steps per day
```{r}
total_spd <- tapply(activity$steps, activity$date,sum)

```
Plotting of the histogram of total number of steps per day
```{r}
hist(total_spd,col="blue",
     xlab="Steps per Day",
     ylab="Frequency", 
     main="Total Steps per day")
```

Computing the Mean of total steps per day
```{r}
mean(total_spd,na.rm=TRUE)
```

Computing the Median of  total steps per day
```{r}
median(total_spd,na.rm=TRUE)
``` 

## What is the average daily activity pattern?
Computing the mean of steps over all days by time interval
```{r}
mean_sbi <- tapply(activity$steps,
                   activity$interval,
                   mean,
                   na.rm=TRUE)
```
Timeseries plot of of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
plot(row.names(mean_sbi),mean_sbi,type="l",
     xlab = "Time Intervals (5-minute)",
     ylab = "Mean number of steps (all Days)",
     main = "Average number of Steps at different 5 minute Intervals",
     col  = "blue")
```
Time interval containing the maximum average number of steps over all days
```{r}
interval_num <- which.max(mean_sbi)
interval_mas <- names(interval_num)
interval_mas
```
## Imputing missing values
Computing the number of NA values in the initial data
```{r}
number_nav <- sum(is.na(activity))
number_nav 
```

Replacing missing values with the average interval value across all days
```{r}
na_indices <- which(is.na(activity))
replaced_values <- mean_sbi[as.character(activity[na_indices,3])]
names(replaced_values) <- na_indices
for (i in na_indices) {
    activity$steps[i] = replaced_values[as.character(i)]
}
sum(is.na(activity)) # Checking that == 0
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
days <- weekdays(activity$date)
activity$day_type <- ifelse(days == "Saturday" | days == "Sunday",
                            "Weekend", "Weekday")
mean_sbi <- aggregate(activity$steps,
                      by=list(activity$interval,
                      activity$day_type),mean)
names(mean_sbi) <- c("interval","day_type","steps")
xyplot(steps~interval | day_type, mean_sbi,type="l",
       layout=c(1,2),
       xlab="Interval",
       ylab = "Steps")

```
Computing mean, median, max and min of the steps across all intervals and days by Weekdays and Weekends
```{r}
tapply(mean_sbi$steps,mean_sbi$day_type,
       function (x) { c(MIN    = min(x),
                        MEAN   = mean(x),
                        MEDIAN = median(x),
                        MAX    = max(x))})
```
